---
title: "Hotel Cancellation - Naive Model"
author: "Caitlin Howansky & Wei Li"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: '3'
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

#__Get Data Ready__
```{r message = FALSE, echo = FALSE, error = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 
# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)
```

```{r message = FALSE, echo = FALSE, error = FALSE}
library(tidyverse) 
library(dplyr) # joins
#library(janitor) # pretty cross-tabs
library(kableExtra) # pretty html tables
library(formattable)
library(gridExtra)
library(scales)
library(pastecs)
library(GGally)
library(lubridate)
library(directlabels)
library(Metrics)
```

```{r message = FALSE, echo = FALSE, error = FALSE}
# disable scientific notation in R
options(scipen=999)
```

```{r message = FALSE, echo = FALSE, error = FALSE}
# load the data files
hot1 <- read_csv('BOOKINGS_ATLCP.csv')
hot2 <- read_csv('BOOKINGS_NYCHA.csv')

```

```{r}
# change date fields to date
hot1$stay_dt <- as.Date(hot1$stay_dt, "%m/%d/%Y")
hot1$booking_dt <- as.Date(hot1$booking_dt, "%m/%d/%Y")
hot2$stay_dt <- as.Date(hot2$stay_dt, "%m/%d/%Y")
hot2$booking_dt <- as.Date(hot2$booking_dt, "%m/%d/%Y")

#create field for non-numerical dow
hot1$day_of_week <- as.factor(hot1$dow)
hot2$day_of_week <- as.factor(hot2$dow)
```

```{r}
# filter OTB=0, truncate the data by dp, add price, otb_cxl_rate, day_type, month, regrouping tags to the dataset
# Reserve last 3 weeks data as Validation Data and the prior are Training Data
hot1 <- hot1 %>% 
  filter(OTB!=0, days_prior<33) %>% 
  mutate(price=OTB_rev/OTB, OTB_cxl_rate = OTB_to_be_cxl/OTB,
         day_type = as.factor(case_when(day_of_week == '7' ~ 'weekend',
                              day_of_week == '1' ~ 'weekend', 
                              TRUE ~'weekday')),
         month = as.factor(month(stay_dt)),
         prod_group_perc = as.factor(case_when(product_type == 'OPAQUE' ~ 'OP/FENC',
                                         product_type == 'FENCED' ~ 'OP/FENC',
                                         product_type == 'CORPORATE' ~ 'CORP/BUS',
                                         product_type == 'BUSINESS TRAVEL AGENCIES' ~ 'CORP/BUS',
                                         product_type == 'GROUP' ~ 'GROUP',
                                         product_type == 'MEMBERSHIP MARKETING' ~ 'MEM/OTHER',
                                         product_type == 'OTHER' ~ 'MEM/OTHER',
                                         product_type == 'UNFENCED' ~ 'GOV/UNFENC',
                                         product_type == 'GOVERNMENT' ~ 'GOV/UNFENC')),
         prod_group_behav = as.factor(case_when(product_type == 'OPAQUE' ~ 'OP/FENC/OTH',
                                         product_type == 'FENCED' ~ 'OP/FENC/OTH',
                                         product_type == 'CORPORATE' ~ 'CORPORATE',
                                         product_type == 'BUSINESS TRAVEL AGENCIES' ~ 'GOV/BUS/MEM',
                                         product_type == 'GROUP' ~ 'GROUP',
                                         product_type == 'MEMBERSHIP MARKETING' ~ 'GOV/BUS/MEM',
                                         product_type == 'OTHER' ~ 'OP/FENC/OTH',
                                         product_type == 'UNFENCED' ~ 'UNFENCED',
                                         product_type == 'GOVERNMENT' ~ 'GOV/BUS/MEM')),
         prod_group_cxl_rate = as.factor(case_when(product_type == 'OPAQUE' ~ 'GOV/UNFEN/OP',
                                         product_type == 'FENCED' ~ 'CORP/FEN/MEM',
                                         product_type == 'CORPORATE' ~ 'CORP/FEN/MEM',
                                         product_type == 'BUSINESS TRAVEL AGENCIES' ~ 'OTH/BTA',
                                         product_type == 'GROUP' ~ 'GROUP',
                                         product_type == 'MEMBERSHIP MARKETING' ~ 'CORP/FEN/MEM',
                                         product_type == 'OTHER' ~ 'OTH/BTA',
                                         product_type == 'UNFENCED' ~ 'GOV/UNFEN/OP',
                                         product_type == 'GOVERNMENT' ~ 'GOV/UNFEN/OP')),
         day_type_cxl_rate = as.factor(case_when(day_of_week=="1"~"SUN", 
                                                 day_of_week=="6"~"FRI/SAT",
                                                 day_of_week=="7"~"FRI/SAT",
                                                 TRUE ~ "MON/TUE/WED/THU")))

hot2 <- hot2 %>% 
  filter(OTB!=0, days_prior<32) %>% 
  mutate(price=OTB_rev/OTB, OTB_cxl_rate = OTB_to_be_cxl/OTB,
         day_type = as.factor(case_when(day_of_week == '7' ~ 'weekend',
                              day_of_week == '1' ~ 'weekend', 
                              TRUE ~'weekday')),
         month = as.factor(month(stay_dt)),
         prod_group_perc = as.factor(case_when(product_type == 'OPAQUE' ~ 'OP/FENC/OTH',
                                         product_type == 'FENCED' ~ 'OP/FENC/OTH',
                                         product_type == 'CORPORATE' ~ 'CORP/GROUP/TACT',
                                         product_type == 'BUSINESS TRAVEL AGENCIES' ~ 'UNF/WHOLE/MEM/BUS',
                                         product_type == 'GROUP' ~ 'CORP/GROUP/TACT',
                                         product_type == 'TACTICAL MARKETING' ~ 'CORP/GROUP/TACT',
                                         product_type == 'MEMBERSHIP MARKETING' ~ 'UNF/WHOLE/MEM/BUS',
                                         product_type == 'OTHER' ~ 'OP/FENC/OTH',
                                         product_type == 'UNFENCED' ~ 'UNF/WHOLE/MEM/BUS',
                                         product_type == 'GOVERNMENT' ~ 'GOVERNMENT',
                                         product_type == 'WHOLESALE' ~ 'UNF/WHOLE/MEM/BUS')),
         prod_group_behav = as.factor(case_when(product_type == 'OPAQUE' ~ 'OTH/OPA/FEN',
                                         product_type == 'FENCED' ~ 'OTH/OPA/FEN',
                                         product_type == 'CORPORATE' ~ 'CORP/MEMB/WHOLE',
                                         product_type == 'BUSINESS TRAVEL AGENCIES' ~ 'BTA/UNFEN/CORP/MM',
                                         product_type == 'GROUP' ~ 'GROUP',
                                         product_type == 'TACTICAL MARKETING' ~ 'TACT/WHOLE',
                                         product_type == 'MEMBERSHIP MARKETING' ~ 'BTA/UNFEN/CORP/MM',
                                         product_type == 'OTHER' ~ 'OTH/OPA/FEN',
                                         product_type == 'UNFENCED' ~ 'BTA/UNFEN/CORP/MM',
                                         product_type == 'GOVERNMENT' ~ 'GOVERNMENT',
                                         product_type == 'WHOLESALE' ~ 'TACT/WHOLE')),
         day_type_cxl_rate = as.factor(case_when(day_of_week=="1"~"SUN", 
                                                 day_of_week=="6"~"FRI/SAT",
                                                 day_of_week=="7"~"FRI/SAT",
                                                 TRUE ~ "MON/TUE/WED/THU")))
h1_train <- hot1[hot1$stay_dt <= (max(hot1$stay_dt) - 21),]
h2_train <- hot2[hot2$stay_dt <= (max(hot2$stay_dt) - 21),]
```


#__Linear Regression Models__
##__Hotel 1__
```{r}
h1_train %>%
  select(days_prior, cummulative_cxl_bookings, cummulative_gross_bookings, cummulative_gross_rev, OTB, OTB_rev, OTB_to_be_cxl, price, daily_gross_bookings, daily_cxl_bookings, daily_net_bookings)%>%
  cor()
```

+ Below variables are highly correlated, so we will not combine them into the same model:
    - cummulative_cxl_bookings, 
    - cummulative_gross_bookings, 
    - cummulative_gross_rev, 
    - OTB,
    - OTB_rev, 
    - OTB_to_be_cxl   

+ Daily_net_bookings shouldn't combine with daily_gross_bookings due to high correlation.     

###__mod1_1: y ~ simple combinations of variables__
```{r}
# Simple combination of variables: days_prior+cummulative_cxl_bookings+daily_gross_bookings+daily_cxl_bookings+price+day_of_week+month+prod_group_perc
mod1_1 <- lm(formula=OTB_cxl_rate~days_prior+cummulative_cxl_bookings+daily_gross_bookings+daily_cxl_bookings+price+day_of_week+month+prod_group_perc, data = h1_train)
summary(mod1_1)
```

+ This combination can show the highest R-square 0.1294. If change cummulative_cxl_bookings to other highly-correlated variables, or change daily_gross_bookings to daily_net_bookings, the results are the same 0.1294. 

+ Tried to use prod_group_behav/prod_group_cxl_rate to replace prod_group_perc, the R-squared is < 0.1, so the prod_group_perc grouping method is the best?

+ Tried to use day_type(weekend/weekday) to replace day_of_week, R-square is 0.126 < 0.1294; 

+ Tried to use day_type_cxl_rate (3 types according to cxl rate curves) to replace day_of_week, r-square is 0.1272 larger than day_type but lower than day_of_week. --- It's for sure, the more categories in a variable, the more information can be captured. Then why we want to do regrouping? Because of small sample size? How to determine whether we need to regroup?

+ There is no need to do logarithms/polynomials on x/y, because of 0 values in the variables and highly right skewed data. 

###__mod1_2: y ~ interaction terms: prod_group_perc*days_prior
```{r}
mod1_2 <- lm(formula=OTB_cxl_rate~days_prior+cummulative_cxl_bookings+daily_gross_bookings+daily_cxl_bookings+price+day_of_week+month+prod_group_perc+prod_group_perc*days_prior, data = h1_train)
summary(mod1_2)
```


###__mod1_3: y ~ interaction terms: prod_group_perc*days_prior + month*days_prior
```{r}
mod1_3 <- lm(formula=OTB_cxl_rate~days_prior+cummulative_cxl_bookings+daily_gross_bookings+daily_cxl_bookings+price+day_of_week+month+prod_group_perc+prod_group_perc*days_prior+month*days_prior, data = h1_train)
summary(mod1_3)
```

+ Here still using day_of_week instead of regrouping of day, because this produces the best R-squared.




##__Hotel 1__
```{r}
h2_train %>%
  select(days_prior, cummulative_cxl_bookings, cummulative_gross_bookings, cummulative_gross_rev, OTB, OTB_rev, OTB_to_be_cxl, price, daily_gross_bookings, daily_cxl_bookings, daily_net_bookings)%>%
  cor()
```

+ Below variables are highly correlated, so we will not combine them into the same model:
    - cummulative_cxl_bookings, 
    - cummulative_gross_bookings, 
    - cummulative_gross_rev, 
    - OTB,
    - OTB_rev, 
    - OTB_to_be_cxl   

+ Daily_net_bookings shouldn't combine with daily_gross_bookings due to high correlation.     



###__mod2_1: y ~ simple combinations of variables__
```{r}
# Simple combination of variables: days_prior+cummulative_cxl_bookings+daily_gross_bookings+daily_cxl_bookings+price+day_of_week+month+prod_group_perc
mod2_1 <- lm(formula=OTB_cxl_rate~days_prior+cummulative_cxl_bookings+daily_gross_bookings+daily_cxl_bookings+price+day_of_week+month+prod_group_perc, data = h2_train)
summary(mod2_1)
```

+ This combination can show the highest R-square 0.1903 If change cummulative_cxl_bookings to other highly-correlated variables, or change daily_gross_bookings to daily_net_bookings, the results are the same 0.1903. 

+ Tried to use prod_group_behav to replace prod_group_perc, the R-squared is < 0.1, so the prod_group_perc grouping method is the best?

+ Tried to use day_type(weekend/weekday) to replace day_of_week, R-square is 0.189 < 0.1903 ; 

+ Tried to use day_type_cxl_rate (3 types according to cxl rate curves) to replace day_of_week, r-square is 0.1895 larger than day_type but lower than day_of_week. --- It's for sure, the more categories in a variable, the more information can be captured. Then why we want to do regrouping? Because of small sample size? How to determine whether we need to regroup?

+ There is no need to do logarithms/polynomials on x/y, because of 0 values in the variables and highly right skewed data. 

###__mod1_2: y ~ interaction terms: prod_group_perc*days_prior
```{r}
mod2_2 <- lm(formula=OTB_cxl_rate~days_prior+cummulative_cxl_bookings+daily_gross_bookings+daily_cxl_bookings+price+day_of_week+month+prod_group_perc+prod_group_perc*days_prior, data = h2_train)
summary(mod2_2)
```


###__mod1_3: y ~ interaction terms: prod_group_perc*days_prior + month*days_prior
```{r}
mod2_3 <- lm(formula=OTB_cxl_rate~days_prior+cummulative_cxl_bookings+daily_gross_bookings+daily_cxl_bookings+price+day_of_week+month+prod_group_perc+prod_group_perc*days_prior+month*days_prior, data = h2_train)
summary(mod2_3)
```

+ Here still using day_of_week instead of regrouping of day, because this produces the best R-squared. 


#__Performance Evaluation__
##__Hotel 1__
###__mod1_1__
```{r}
h1_test <- hot1[hot1$stay_dt > (max(hot1$stay_dt) - 21),]

h1_test$predict_cxl_rate <-  predict(mod1_1, h1_test, se.fit=TRUE)$fit
h1_test$fcst_svv <- h1_test$OTB - h1_test$OTB * h1_test$predict_cxl_rate

```

```{r}
# MAE
mod1_1_mae <- h1_test %>% 
  filter(days_prior!=0) %>% 
  mutate(dp_range = case_when(days_prior >=1 & days_prior<=7 ~ '1_7',
                                  days_prior >=8 & days_prior<=14 ~ '8_14',
                                  days_prior >=15 & days_prior<=21 ~ '15_21',
                                  days_prior >=22 & days_prior<=28 ~ '22_28',
                                  days_prior >=29 ~ '29_32')) %>% 
  group_by(dp_range) %>% 
  summarise(meanAE = mae(OTB_to_survive, fcst_svv))
mod1_1_mae
```

```{r}
# MAPE: the OTB_to_survive in GOVERNMENT contains 0 values, so MAPE is INF
mod1_1_mape <- h1_test %>%  
  filter(days_prior!=0, OTB_to_survive!=0) %>% 
  mutate(dp_range = case_when(days_prior >=1 & days_prior<=7 ~ '1_7',
                                  days_prior >=8 & days_prior<=14 ~ '8_14',
                                  days_prior >=15 & days_prior<=21 ~ '15_21',
                                  days_prior >=22 & days_prior<=28 ~ '22_28',
                                  days_prior >=29 ~ '29_32')) %>% 
  group_by(dp_range) %>% 
  summarise(meanAPE = mape(OTB_to_survive, fcst_svv))
mod1_1_mape
```

```{r}
# MASE
## Avg survival rate:
avg_svv_rt <- sum(h1_train$OTB_to_survive)/sum(h1_train$OTB)

mod1_1_mase <- h1_test %>%  
  filter(days_prior!=0) %>% 
  mutate(dp_range = case_when(days_prior >=1 & days_prior<=7 ~ '1_7',
                                  days_prior >=8 & days_prior<=14 ~ '8_14',
                                  days_prior >=15 & days_prior<=21 ~ '15_21',
                                  days_prior >=22 & days_prior<=28 ~ '22_28',
                                  days_prior >=29 ~ '29_32'),
         abs_diff_1=abs(OTB_to_survive-fcst_svv),
         abs_diff_2=abs(OTB_to_survive-avg_svv_rt*OTB)) %>% 
  group_by(dp_range) %>% 
  summarise(meanASE = sum(abs_diff_1)/sum(abs_diff_2))
mod1_1_mase
```

```{r}
# MAE, MAPE, MASE for mod1_1
left_join(left_join(mod1_1_mae,mod1_1_mape, by=c("dp_range")), mod1_1_mase, by=c("dp_range"))
```


