---
title: "Hotel Cancellation - Naive Model"
author: "Caitlin Howansky & Wei Li"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: '3'
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---
```{r, include=FALSE}
#run Naive_Model_final first
load("my_work_space.RData")

```



##__Atlanta Hotel__
```{r, echo=FALSE}
hotel1
hotel1_in_sample
```

During this Naive Model Analysis, we found that our second way of grouping the product types was not as good as predicting as our other grouping methods are. In the table above, the column labeled days prior only takes into account the number of days a customer booked before their stay date. Group 1 above is our method of grouping that took the closest cancellation rates and grouped them together while Group 2 looked at the cancellation curve. DOW references the grouping of week days by specific day of week and Day Type groups the days of the we week into week days and weekends. When grouping by group 1, it seems as though grouping it with days of week has a better prediction than grouping it with the type of day. However, when grouping with our third group, the type of day has a slightly better error rate. The best group method for the Atlanta Hotel is when using the group based on the cancellation curve and type of day. For most of the methods, the error is lowest for the first two groupings of days prior and then gets worse for days prior between 15 and 28, but then gets a little better. This means that the models are better at predicting cancellation rates when they are closer to the stay date. 

##__NY Hotel__
```{r, echo=FALSE}
hotel2
hotel2_in_sample
```

The New York Hotel has very different numbers than the Atlanta Hotel. It looks like the MAE numbers are much worse but the MAPE numbers are better or at least comparable. Group 1 in the table above shows the groupings based on cancellation rates while group 2 is based on cancellation curves.  For this hotel, our grouping methods show that there is usually a direct relationship between error and days prior. Comparing Group 1 and DOW and Group 2 and DOW, Group 2 out performs Group 1 in every grouping of days prior. However, when using group 2 Group 1 out performs Group 2 in days prior 1-7 and 29-32 according to the MAPE and MASE numbers. For this hotel, like Atlanta, the best overall grouping is the cancellation curve product type grouping along with the type of day it is. 
