---
title: "Capstone EDA - Caitlin"
author: "Caitlin Howansky"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: '3'
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r message = FALSE, echo = FALSE, error = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 
# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)
```

```{r message = FALSE, echo = FALSE, error = FALSE}
library(tidyverse) 
library(dplyr) # joins
#library(janitor) # pretty cross-tabs
library(kableExtra) # pretty html tables
library(formattable)
library(gridExtra)
library(scales)
library(pastecs)
library(GGally)
library(lubridate)
```



```{r message = FALSE, echo = FALSE, error = FALSE}
# disable scientific notation in R
options(scipen=999)
```

```{r}
# load the data files
hot1 <- read_csv('BOOKINGS_ATLCP.csv')
hot2 <- read_csv('BOOKINGS_NYCHA.csv')

```

##__EDA__

##__Data Overview__
```{r}
#first look at all tables
head(hot1)
head(hot2)
```

```{r}
#summaries of data
summary(hot1)
summary(hot2)
```
__ Hotel 1
- no nulls
- days of week are numerical - rename?
- booking date - character not date field
- min of -4844 for daily net revenue 
- should we group product_type?

__ Hotel 2
- similar to hotel 


```{r}
# change date fields to date
hot1$stay_dt <- as.Date(hot1$stay_dt, "%m/%d/%Y")
hot1$booking_dt <- as.Date(hot1$booking_dt, "%m/%d/%Y")
hot2$stay_dt <- as.Date(hot2$stay_dt, "%m/%d/%Y")
hot2$booking_dt <- as.Date(hot2$booking_dt, "%m/%d/%Y")

summary(hot1)

#create field for non-numerical dow
hot1$day_of_week <- as.factor(hot1$dow)
hot2$day_of_week <- as.factor(hot2$dow)
```

```{r}
# Reserve last 3 weeks data as Validation Data and the prior are Training Data
h1_train <- hot1[hot1$stay_dt <= (max(hot1$stay_dt) - 21),]
h2_train <- hot2[hot2$stay_dt <= (max(hot2$stay_dt) - 21),]
```

##__Hotel 1__

```{r}
#look into product type
h1_train %>%
  select(product_type,  OTB, OTB_to_be_cxl) %>%
  group_by(product_type) %>%
  summarise(avg_bookings = mean(OTB),
            avg_cancels = mean(OTB_to_be_cxl),
            perc_cxl = (mean(OTB_to_be_cxl)/mean(OTB))*100) %>%
  kable(align = c("c", "c", "c")) %>% 
  kable_styling(bootstrap_options = c("striped", "condensed", "bordered")) %>% 
  column_spec(1:4, color = "#000000")
```
- group has a significantly higher average otb bookings and cancellations
- corporate is the second highest for both avg bookings and cancels
- fenced product type seems to help decrease the number of cancellations 
- highest percentage of cancellations is other then government 
- lowest percentageis opaque and fenced
- group by percentage calcs?
      - (opaque, fenced), (corporate,business travel agencies), group, (membership, other), (unfenced,             government)




```{r}
#above by day of week
h1_train %>%
  select(product_type, day_of_week, OTB, OTB_to_be_cxl) %>%
  group_by(product_type, day_of_week) %>%
  summarise(avg_bookings = mean(OTB),
            avg_cancels = mean(OTB_to_be_cxl)) %>%
  arrange(day_of_week) %>%
  kable(align = c("c", "c", "c", "c")) %>% 
  kable_styling(bootstrap_options = c("striped", "condensed", "bordered")) %>% 
  column_spec(1:4, color = "#000000")
```


```{r}
#look at bookings by day of week and product type

h1_train %>%
  select(day_of_week, product_type, OTB) %>%
  group_by(day_of_week, product_type) %>%
  summarise(avg_bookings = mean(OTB)) %>%
  ggplot(aes(x = reorder(day_of_week, -avg_bookings), y = avg_bookings, fill = product_type)) +
           geom_bar(stat = "identity", width = .8, color = "white", position = "dodge") +
           theme_minimal() +
           theme(axis.title = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.text.y = element_blank(),
                axis.text.x = element_text(size = 11, face = "bold")) + 
           ggtitle("Average # of Bookings by DOW and Product Type")
```
- dow 1 is lowest for all but corporate
- group and corporate have a very different weekly pattern (group separately)
- other accounts for a very small portion 
- business travel angencies are also very low (lowest on friday and saturday)
- day 1 is sunday I looked it up

```{r}
#above by cancellations
h1_train %>%
  select(day_of_week, product_type, OTB_to_be_cxl) %>%
  group_by(day_of_week, product_type) %>%
  summarise(avg_cxl = mean(OTB_to_be_cxl)) %>%
  ggplot(aes(x = reorder(day_of_week, -avg_cxl), y = avg_cxl, fill = product_type)) +
           geom_bar(stat = "identity", width = .8, color = "white", position = "dodge") +
           theme_minimal() +
           theme(axis.title = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.text.y = element_text(size = 11, face = "bold"),
                axis.text.x = element_text(size = 11, face = "bold")) + 
           ggtitle("Average # of CXLs by DOW and Product Type")
```

- Another grouping possibility I'm thinking about: Group, Corporate, unfenced, (other, opaque, fenced), (bus, gov, mem) totalling 5 goups 

- day 6 sees most cancellations and day 1 sees least

```{r}
#revenue by day of week

h1_train %>%
  select(day_of_week, product_type, OTB_rev) %>%
  group_by(day_of_week, product_type) %>%
  summarise(avg_rev = mean(OTB_rev)) %>%
  ggplot(aes(x = reorder(day_of_week, -avg_rev), y = avg_rev, fill = product_type)) +
           geom_bar(stat = "identity", width = .8, color = "white", position = "dodge") +
           theme_minimal() +
           theme(axis.title = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.text.y = element_text(size = 11, face = "bold"),
                axis.text.x = element_text(size = 11, face = "bold")) + 
           ggtitle("Avg Revenues by DOW and product type")
```

```{r}
#look into days prior 

h1_train %>%
  select(OTB, OTB_to_be_cxl, OTB_rev, days_prior) %>%
  group_by(days_prior) %>%
  summarise(avg_bookings = mean(OTB),
            avg_cancels = mean(OTB_to_be_cxl),
            avg_otb_rev = mean(OTB_rev)) %>%
  arrange(desc(avg_otb_rev)) %>%
  kable(align = c("c", "c", "c", "c")) %>% 
  kable_styling(bootstrap_options = c("striped", "condensed", "bordered")) %>% 
  column_spec(1:4, color = "#000000")
```

- a lot of bookings  happen the day of. More bookings than cancellations creating higher net revenues
- highest net revenues are either very close to the day of staying or around 20 days out
- truncate between days 24 and 30 

```{r}
#graphical representation of above

h1_train %>%
  select(days_prior, OTB, OTB_to_be_cxl, OTB_rev) %>%
  group_by(days_prior) %>%
  summarise(avg_net_rev = mean(OTB_rev),
            avg_bookings = mean(OTB),
            avg_cxl = mean(OTB_to_be_cxl)) %>%
  ggplot() +
  geom_line(aes(x = days_prior, y = avg_bookings), color = "red") +
  geom_line(aes(x = days_prior, y = avg_cxl), color = "blue") +
           theme_minimal() +
           theme(axis.title = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.text.y = element_text(size = 11, face = "bold"),
                axis.text.x = element_text(size = 11, face = "bold")) + 
           ggtitle("Bookings & Cancellations by Days Prior")
```

- In the article we read, we read that cancellations are like a boat structure meaning that there are more cancellations and bookings awhile before and right before. If this is the case, the other side of the boat would be many more days out than 60 


```{r}
h1_train %>%
  select(days_prior, cummulative_gross_bookings, product_type) %>%
  group_by(days_prior, product_type) %>%
  summarise(avg_bookings = mean(cummulative_gross_bookings)) %>%
  ggplot() +
  geom_line(aes(x = days_prior, y = avg_bookings,  color = product_type)) +
           theme_minimal() +
           theme(axis.title = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.text.y = element_text(size = 11, face = "bold"),
                axis.text.x = element_text(size = 11, face = "bold")) + 
           ggtitle("Cumm Bookings by Product Type and Days Prior")
```

- no big bumps in cummulative gross bookings - following a logical inverse relationship

- Are there any outliers here?
- using daily gross bookings to see if there is a huge spike in demand

```{r}
h1_train %>%
  select(days_prior, daily_gross_bookings, product_type) %>%
  group_by(days_prior, product_type) %>%
  summarise(avg_bookings = mean(daily_gross_bookings)) %>%
  ggplot() +
  geom_line(aes(x = days_prior, y = avg_bookings,  color = product_type)) +
           theme_minimal() +
           theme(axis.title = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.text.y = element_text(size = 11, face = "bold"),
                axis.text.x = element_text(size = 11, face = "bold")) + 
           ggtitle("Bookings by Product Tpe and Days Prior")
```
- group has huge spike around day 22 - outlier?
- group has big decrease around day 11


```{r}
#looking into outliers
group_outlier <- h1_train %>%
                filter(product_type == 'GROUP')

group_outlier <- boxplot(group_outlier$daily_gross_bookings)$out

h1_train %>%
  select(product_type, days_prior, daily_gross_bookings) %>%
  filter(product_type == 'GROUP') %>%
  filter(daily_gross_bookings > 200)


corp_outlier <- h1_train %>%
                filter(product_type == 'CORPORATE')

corp_outlier <- boxplot(corp_outlier$daily_gross_bookings)$out

h1_train %>%
  select(product_type, days_prior, daily_gross_bookings) %>%
  filter(product_type == 'CORPORATE') %>%
  filter(daily_gross_bookings > 20)

gov_outlier <- h1_train %>%
                filter(product_type == 'GOVERNMENT')

gov_outlier <- boxplot(gov_outlier$daily_gross_bookings)$out

h1_train %>%
  select(product_type, days_prior, daily_gross_bookings) %>%
  filter(product_type == 'GOVERNMENT') %>%
  filter(daily_gross_bookings > 10)
```
- 451 outliers for group out of 4148 
    - cut out greater than 400?
    - there are four greater than 400 at days prior 22, 23, 24,25
    - next lowest one is 290

- 421 outliers for corporate 
    - cut out greater than 35?
    - four greater than 35 at 0 days prior 
    - next lowest one is 29
    
- 377 outliers for gov out of 4148 
    - four rows greater than 10 at 16, 23, 25, 23
    

    
    
- what about cancellations?

```{r}
group_outlier <- h1_train %>%
                filter(product_type == 'GROUP')

group_outlier <- boxplot(group_outlier$daily_cxl_bookings)$out

h1_train %>%
  select(product_type, days_prior, daily_cxl_bookings) %>%
  filter(product_type == 'GROUP') %>%
  filter(daily_cxl_bookings > 40)


corp_outlier <- h1_train %>%
                filter(product_type == 'CORPORATE')

corp_outlier <- boxplot(corp_outlier$daily_cxl_bookings)$out

h1_train %>%
  select(product_type, days_prior, daily_cxl_bookings) %>%
  filter(product_type == 'CORPORATE') %>%
  filter(daily_cxl_bookings > 20)

gov_outlier <- h1_train %>%
                filter(product_type == 'GOVERNMENT')

gov_outlier <- boxplot(gov_outlier$daily_cxl_bookings)$out

h1_train %>%
  select(product_type, days_prior, daily_cxl_bookings) %>%
  filter(product_type == 'GOVERNMENT') %>%
  filter(daily_cxl_bookings > 10)
```


```{r}
h1_train %>%
  select(days_prior, daily_cxl_bookings, product_type) %>%
  group_by(days_prior, product_type) %>%
  summarise(avg_bookings = mean(daily_cxl_bookings)) %>%
  ggplot() +
  geom_line(aes(x = days_prior, y = avg_bookings, fill = product_type, color = product_type)) +
           theme_minimal() +
           theme(axis.title = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.text.y = element_text(size = 11, face = "bold"),
                axis.text.x = element_text(size = 11, face = "bold")) + 
           ggtitle("Cancellations by Product Tpe and Days Prior")
```
- no huge spikes in cancellations 




```{r}
#net revenue and days prior

h1_train %>%
  select(days_prior, OTB, OTB_to_be_cxl, OTB_rev) %>%
  group_by(days_prior) %>%
  summarise(avg_net_rev = mean(OTB_rev)) %>%
  ggplot() +
  geom_line(aes(x = days_prior, y = avg_net_rev, color = "red")) +
           theme_minimal() +
           theme(axis.title = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.text.y = element_text(size = 11, face = "bold"),
                axis.text.x = element_text(size = 11, face = "bold")) + 
           ggtitle("OTB Revenue by Days Prior")

```
 - net revenue has same bumps as bookings since the cancellation line is much smoother


- Does month change bookings or cancellations? 

```{r}
h1_train$month = month(h1_train$stay_dt)

h1_train %>%
  select(month, product_type, OTB) %>%
  group_by(month, product_type) %>%
  summarise(avg_bookings = mean(OTB)) %>%
  ggplot(aes(x = reorder(month, -avg_bookings), y = avg_bookings, fill = product_type)) +
           geom_bar(stat = "identity", width = .8, color = "white", position = "dodge") +
           theme_minimal() +
           theme(axis.title = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.text.y = element_text(size = 11, face = "bold"),
                axis.text.x = element_text(size = 11, face = "bold")) + 
           ggtitle("Average # of Bookings by Month and Product Type")
```

- hard to tell with only three months but staying in april is significantly less than february and march for group
- more opaque in april because they probably sold more last minute






##__ Hotel 2 __
```{r}
#look into product type
h2_train %>%
  select(product_type,  OTB, OTB_to_be_cxl) %>%
  group_by(product_type) %>%
  summarise(avg_bookings = mean(OTB),
            avg_cancels = mean(OTB_to_be_cxl),
             perc_cxl = (mean(OTB_to_be_cxl)/mean(OTB))*100) %>%
  kable(align = c("c", "c", "c")) %>% 
  kable_styling(bootstrap_options = c("striped", "condensed", "bordered")) %>% 
  column_spec(1:3, color = "#000000")
```
- The NY hotel has completely different average bookings for product types
- unfenced, opaque, group, & corporate all have an average of between 3.3 and 3.7 bookings a day
- group product types differently than hotel 1
- highest percent cancellations are wholesale and government 
- lowest is other and fenced
- group by percentage calcs?


```{r}
#above by day of week
h2_train %>%
  select(product_type, day_of_week, OTB, OTB_to_be_cxl) %>%
  group_by(product_type, day_of_week) %>%
  summarise(avg_bookings = mean(OTB),
            avg_cancels = mean(OTB_to_be_cxl)) %>%
  arrange(day_of_week) %>%
  kable(align = c("c", "c", "c", "c")) %>% 
  kable_styling(bootstrap_options = c("striped", "condensed", "bordered")) %>% 
  column_spec(1:4, color = "#000000")
```

```{r}
#bookings by day of week and product type

h2_train %>%
  select(day_of_week, product_type, OTB) %>%
  group_by(day_of_week, product_type) %>%
  summarise(avg_bookings = mean(OTB)) %>%
  ggplot(aes(x = reorder(day_of_week, -avg_bookings), y = avg_bookings, fill = product_type)) +
           geom_bar(stat = "identity", width = .8, color = "white", position = "dodge") +
           theme_minimal() +
           theme(axis.title = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.text.y = element_text(size = 11, face = "bold"),
                axis.text.x = element_text(size = 11, face = "bold")) + 
           ggtitle("Average # of Bookings by DOW and Product Type")
```
 - opaque is a very high category especially on days 6/7 (friday, saturday stay dates)
 - other, government, tactical marketing very small categories
 - seems like this hotel houses a lot of business since corporate and group are so high and they dont have as many staying on fridays and saturday nights 
 - unfenced remains relatively the same 


```{r}
#above by cancellations
h2_train %>%
  select(day_of_week, product_type, OTB_to_be_cxl) %>%
  group_by(day_of_week, product_type) %>%
  summarise(avg_cxl = mean(OTB_to_be_cxl)) %>%
  ggplot(aes(x = reorder(day_of_week, -avg_cxl), y = avg_cxl, fill = product_type)) +
           geom_bar(stat = "identity", width = .8, color = "white", position = "dodge") +
           theme_minimal() +
           theme(axis.title = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.text.y = element_text(size = 11, face = "bold"),
                axis.text.x = element_text(size = 11, face = "bold")) + 
           ggtitle("Average # of CXLs by DOW and Product Type")
```

- corporate highest to ccancel
- unfenced cancels a lot as well

```{r}
#revenue by day of week

h2_train %>%
  select(day_of_week, product_type, OTB_rev) %>%
  group_by(day_of_week, product_type) %>%
  summarise(avg_rev = mean(OTB_rev)) %>%
  ggplot(aes(x = reorder(day_of_week, -avg_rev), y = avg_rev, fill = product_type)) +
           geom_bar(stat = "identity", width = .8, color = "white", position = "dodge") +
           theme_minimal() +
           theme(axis.title = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.text.y = element_text(size = 11, face = "bold"),
                axis.text.x = element_text(size = 11, face = "bold")) + 
           ggtitle("Avg Revenues by DOW and product type")
```
-sunday has lowest revenues

########look into price - rev/#bookings    look at correlation between price and cancelation rate#####

```{r}
#look into days prior 

h2_train %>%
  select(OTB, OTB_to_be_cxl, OTB_rev, days_prior) %>%
  group_by(days_prior) %>%
  summarise(avg_bookings = mean(OTB),
            avg_cancels = mean(OTB_to_be_cxl),
            avg_net_rev = mean(OTB_rev)) %>%
  arrange(days_prior) %>%
  kable(align = c("c", "c", "c", "c")) %>% 
  kable_styling(bootstrap_options = c("striped", "condensed", "bordered")) %>% 
  column_spec(1:4, color = "#000000")
```

- similar to revenue distribution of hotel 1 
- most people tend to book about a week out
- look into grouping by weeks out?


```{r}
#graphical representation of above

h2_train %>%
  select(days_prior, OTB, OTB_to_be_cxl, OTB_rev) %>%
  group_by(days_prior) %>%
  summarise(avg_net_rev = mean(OTB_rev),
            avg_bookings = mean(OTB),
            avg_cxl = mean(OTB_to_be_cxl)) %>%
  ggplot() +
  geom_line(aes(x = days_prior, y = avg_bookings), color = "red") +
  geom_line(aes(x = days_prior, y = avg_cxl), color = "blue") +
           theme_minimal() +
           theme(axis.title = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.text.y = element_text(size = 11, face = "bold"),
                axis.text.x = element_text(size = 11, face = "bold")) + 
           ggtitle("Bookings & Cancellations by Days Prior")
```


```{r}
#net revenue and days prior

h2_train %>%
  select(days_prior, OTB, OTB_to_be_cxl, OTB_rev) %>%
  group_by(days_prior) %>%
  summarise(avg_net_rev = mean(OTB_rev)) %>%
  ggplot() +
  geom_line(aes(x = days_prior, y = avg_net_rev, color = "red")) +
           theme_minimal() +
           theme(axis.title = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.text.y = element_text(size = 11, face = "bold"),
                axis.text.x = element_text(size = 11, face = "bold")) + 
           ggtitle("Bookings & Cancellations by Days Prior")

```

- steep hills between days 0 and 20


```{r}
h2_train$month = month(h2_train$stay_dt)

h2_train %>%
  select(month, product_type, OTB) %>%
  group_by(month, product_type) %>%
  summarise(avg_bookings = mean(OTB)) %>%
  ggplot(aes(x = reorder(month, -avg_bookings), y = avg_bookings, fill = product_type)) +
           geom_bar(stat = "identity", width = .8, color = "white", position = "dodge") +
           theme_minimal() +
           theme(axis.title = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.text.y = element_text(size = 11, face = "bold"),
                axis.text.x = element_text(size = 11, face = "bold")) + 
           ggtitle("Average # of Bookings by Month and Product Type")
```

- very different from hotel 1
- reasonablly the same for each month

