---
title: "Wei - EDA"
author: "Wei"
date: "10/01/2019"
output:
  html_document:
    df_print: paged
---


```{r include = FALSE}
# Course: BUAN 5510
# Purpose: EDA
# Author: Wei Li  
# Date: 10/01/2019

```

```{r message = FALSE, echo = FALSE, error = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)

```

```{r message = FALSE, echo = FALSE, error = FALSE}
library(tidyverse) 
library(dplyr) # joins
#library(janitor) # pretty cross-tabs
library(kableExtra) # pretty html tables
library(formattable)
library(gridExtra)
library(scales)
library(pastecs)
library(GGally)
library(lubridate)
```


```{r}
# Load 2 hotel data
hot1 <- read.csv('../raw_data/BOOKINGS_ATLCP.csv')
hot2 <- read.csv('../raw_data/BOOKINGS_NYCHA.csv')
```

```{r}
# EDA only on training data 
## Change date to date type
hot1$stay_dt <- as.Date(hot1$stay_dt, "%m/%d/%Y")
hot1$booking_dt <- as.Date(hot1$booking_dt, "%m/%d/%Y")
hot2$stay_dt <- as.Date(hot2$stay_dt, "%m/%d/%Y")
hot2$booking_dt <- as.Date(hot2$booking_dt, "%m/%d/%Y")

#create field for non-numerical dow
hot1$day_of_week <- as.factor(hot1$dow)
hot2$day_of_week <- as.factor(hot2$dow)

## Reserve last 3 weeks data as Validation Data and the prior are Training Data
h1_train <- hot1[hot1$stay_dt <= (max(hot1$stay_dt) - 21),]
h2_train <- hot2[hot2$stay_dt <= (max(hot2$stay_dt) - 21),]

```

##__Hotel 1__

```{r}
head(h1_train)
```

```{r}
# number of bookings and cancels in each days_prior
h1_train %>% 
  select(days_prior, daily_gross_bookings, daily_cxl_bookings) %>% 
  group_by(days_prior) %>% 
  summarise(avg_bookings = mean(daily_gross_bookings),
            avg_cancels = mean(daily_cxl_bookings),
            sum_bookings = sum(daily_gross_bookings),
            sum_cancels = sum(daily_cxl_bookings)) %>% 
  arrange(avg_cancels)
```

- Above shows the avg and sum of bookings in terms of each days_prior. If we want to truncate, what about days_prior=24 where the avg_cancels = 0.09? In Petraru's paper, he used 16 and 22 in some graphs. So I guess we could choose a close one?

```{r}
# number of cancellations in each days_prior/each product type
h1_train %>% 
  select(days_prior, daily_cxl_bookings) %>% 
  group_by(days_prior) %>% 
  summarize(num = sum(daily_cxl_bookings))

h1_train %>% 
  select(product_type, daily_cxl_bookings) %>% 
  group_by(product_type) %>% 
  summarize(num = sum(daily_cxl_bookings))
```

- days_prior about 25, the number of cxl is about 50
- product_type: need to regroup, due to the few cxls in some types. Regroup method, close cxl rate --> group together. Plot the curves would help. 

```{r}
# cxl rate curve = cummulative_cxl_bookings/cummulative_gross_bookings)
h1_train %>% 
  select(product_type, days_prior, cummulative_cxl_bookings, cummulative_gross_bookings) %>% 
  mutate(cxl_rate = cummulative_cxl_bookings/cummulative_gross_bookings) %>% 
  group_by(days_prior, product_type) %>% 
  summarise(avg_cxl_rate = mean(cxl_rate)) %>% 
  ggplot()+
  geom_line(aes(x=days_prior, y=avg_cxl_rate, color = product_type))

```

Possibly group:
-group 1: Unfenced, membership marketing,
-group 2: fenced
-group 3: corporate, group, business travel
-group 4: opaque

-government and other are not here?

```{r}
# 1-1. Understand the patterns of cancellation rate curves. Try to see if cancellation rates are different by DOW (day of week), product types (by different levels),  month, OTB (demand volume), daily/cumulative gross bookings, daily/cumulative cancellations, price, and etc. Graphical display (plots) will be helpful to understand the patterns.

# DOW
h1_train %>% 
  select(day_of_week, daily_cxl_bookings, daily_gross_bookings) %>% 
  group_by(day_of_week) %>% 
  summarise(avg_bookings = mean(daily_gross_bookings),
            avg_cancels = mean(daily_cxl_bookings),
            sum_bookings = sum(daily_gross_bookings),
            sum_cancels = sum(daily_cxl_bookings)) %>%
  arrange(day_of_week) %>%
  kable(align = c("c", "c", "c", "c")) %>% 
  kable_styling(bootstrap_options = c("striped", "condensed", "bordered")) %>% 
  column_spec(1:3, color = "#000000")

```

- day_of_week 1, 6, 7: lower avg_bookings and lower avg_cancels. Does this make sense? Shouldn't Sunday, Friday, Saturday be the most popular days for hotel? Why the bookings are lower than other days? --- make sense, many group traveler are on business travel during in week days. 
- regroup into 2345 and 167 2 groups? - Nope.

```{r}
# product types
h1_train %>% 
  select(product_type, daily_cxl_bookings, daily_gross_bookings) %>% 
  group_by(product_type) %>% 
  summarise(avg_bookings = mean(daily_gross_bookings), 
            avg_cancels = mean(daily_cxl_bookings),
            perc_cxl = avg_cancels/avg_bookings) %>%
  arrange(perc_cxl) %>%
  kable(align = c("c", "c", "c", "c")) %>% 
  kable_styling(bootstrap_options = c("striped", "condensed", "bordered")) %>% 
  column_spec(1:3, color = "#000000")

```

- Opaque, fenced, other - too few observations(cancels) to use, consider exclude them?

```{r}
# month 
h1_train$month = month(h1_train$stay_dt)

h1_train %>%
  select(month, daily_gross_bookings) %>%
  group_by(month) %>%
  summarise(avg_bookings = mean(daily_gross_bookings)) %>%
  ggplot(aes(x = reorder(month, -avg_bookings), y = avg_bookings)) +
           geom_bar(stat = "identity", position = "dodge") +
           theme_minimal() +
           theme(axis.title = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.text.y = element_text(size = 11, face = "bold"),
                axis.text.x = element_text(size = 11, face = "bold")) + 
           ggtitle("Average # of Bookings by Month")

```

- Avg # in April is lower, only 3 months -- too short to tell, so we could try week, like below.

```{r}
# week 
h1_train$week = week(h1_train$stay_dt)

h1_train %>%
  select(week, daily_gross_bookings) %>%
  group_by(week) %>%
  summarise(avg_bookings = mean(daily_gross_bookings)) %>%
  ggplot(aes(x = week, y = avg_bookings)) +
           geom_bar(stat = "identity", position = "dodge") +
           theme_minimal() +
           theme(axis.title = element_blank(),
                axis.line.y = element_blank(),
                axis.ticks.y = element_blank(),
                axis.text.y = element_text(size = 11, face = "bold"),
                axis.text.x = element_text(size = 11, face = "bold")) + 
           ggtitle("Average # of Bookings by week")
```

- Trend: going down.

```{r}
# OTB (demand volume)

```

```{r}
# daily/cumulative gross bookings, daily/cumulative cancellations, price, and etc. 
```

