---
title: "Hotel Cancellation - KNN"
author: "Caitlin Howansky & Wei Li"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: '3'
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

#__Get Data Ready__
```{r message = FALSE, echo = FALSE, error = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 
# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)
```

```{r message = FALSE, echo = FALSE, error = FALSE}
library(tidyverse) 
library(dplyr) # joins
#library(janitor) # pretty cross-tabs
library(kableExtra) # pretty html tables
library(formattable)
library(gridExtra)
library(scales)
library(pastecs)
library(GGally)
library(lubridate)
library(directlabels)
library(Metrics)
library(FNN)
library(MASS)
```

```{r message = FALSE, echo = FALSE, error = FALSE}
# disable scientific notation in R
options(scipen=999)
```

```{r message = FALSE, echo = FALSE, error = FALSE}
# load the data files
hot1 <- read_csv('../BOOKINGS_ATLCP.csv')
hot2 <- read_csv('../BOOKINGS_NYCHA.csv')
```

```{r}
# change date fields to date
hot1$stay_dt <- as.Date(hot1$stay_dt, "%m/%d/%Y")
hot1$booking_dt <- as.Date(hot1$booking_dt, "%m/%d/%Y")
hot2$stay_dt <- as.Date(hot2$stay_dt, "%m/%d/%Y")
hot2$booking_dt <- as.Date(hot2$booking_dt, "%m/%d/%Y")

#create field for non-numerical dow
hot1$day_of_week <- as.factor(hot1$dow)
hot2$day_of_week <- as.factor(hot2$dow)

```

```{r}
# Function to get MAE, MAPE, MASE
eval <- function(test){
  test$fcst_svv <- test$OTB - test$OTB * test$predict_cxl_rate
  tst_mae <- test %>% 
    filter(days_prior!=0) %>% 
    mutate(dp_range = case_when(days_prior >=1 & days_prior<=7 ~ '1_7',
                                  days_prior >=8 & days_prior<=14 ~ '8_14',
                                  days_prior >=15 & days_prior<=21 ~ '15_21',
                                  days_prior >=22 & days_prior<=28 ~ '22_28',
                                  days_prior >=29 ~ '29_32')) %>% 
    group_by(dp_range) %>% 
    summarise(meanAE = mae(OTB_to_survive, fcst_svv))
  tst_mape <- test %>%  
    filter(OTB_to_survive!=0, days_prior!=0) %>% 
    mutate(dp_range = case_when(days_prior >=1 & days_prior<=7 ~ '1_7',
                                  days_prior >=8 & days_prior<=14 ~ '8_14',
                                  days_prior >=15 & days_prior<=21 ~ '15_21',
                                  days_prior >=22 & days_prior<=28 ~ '22_28',
                                  days_prior >=29 ~ '29_32')) %>% 
    group_by(dp_range) %>% 
    summarise(meanAPE = mape(OTB_to_survive, fcst_svv))
  tst_mase <- test %>%  
    filter(days_prior!=0) %>% 
    mutate(dp_range = case_when(days_prior >=1 & days_prior<=7 ~ '1_7',
                                  days_prior >=8 & days_prior<=14 ~ '8_14',
                                  days_prior >=15 & days_prior<=21 ~ '15_21',
                                  days_prior >=22 & days_prior<=28 ~ '22_28',
                                  days_prior >=29 ~ '29_32'),
         abs_diff_1=abs(OTB_to_survive-fcst_svv),
         abs_diff_2=abs(OTB_to_survive-avg_svv_rt*OTB)) %>% 
    group_by(dp_range) %>% 
    summarise(meanASE = sum(abs_diff_1)/sum(abs_diff_2))
  left_join(left_join(tst_mae,tst_mape, by=c("dp_range")), tst_mase, by=c("dp_range"))
}

```

#__Hotel 1__
```{r}
summary(hot1)
```


##__Regroup day_of_week & product_type into 2-3 groups__
```{r}
# day_of_week cxl curves
hot1 %>% 
  filter(OTB!=0, days_prior<33, stay_dt <= max(stay_dt - 21)) %>% 
  group_by(days_prior, day_of_week) %>% 
  summarise(avg_cxl_rate = mean(OTB_to_be_cxl)/mean(OTB)) %>% 
  ggplot()+
  geom_line(aes(x=days_prior, y=avg_cxl_rate, color = day_of_week))+
  geom_dl(aes(label = day_of_week, x=days_prior, y=avg_cxl_rate, color = day_of_week), method = list(dl.combine("first.points", "last.points"), cex=0.5))

# product_type cxl curves
hot1 %>% 
  filter(OTB!=0, days_prior<33, stay_dt <= max(stay_dt - 21)) %>% 
  group_by(days_prior, product_type) %>% 
  summarise(avg_cxl_rate = mean(OTB_to_be_cxl)/mean(OTB)) %>% 
  ggplot()+
  geom_line(aes(x=days_prior, y=avg_cxl_rate, color = product_type))+
  geom_dl(aes(label = product_type, x=days_prior, y=avg_cxl_rate, color = product_type), method = list(dl.combine("first.points", "last.points"), cex=0.5))
```

+ According to the cxl curve, regroup the 7 days into 2 groups: 1,6,7 and 2,3,4,5

+ Regroup the product types into 2 groups:
    
    - other, business travel agencies, membership marketing
    - corporate, unfenced, group, government, fenced, opaque
    
```{r}
# filter OTB=0, truncate the data by dp, add price, otb_cxl_rate, regrp_type, month, regrouping tags to the dataset
# day_type = 1: 'FRI/SAT/SUN'; = 0:'MON/TUE/WED/THU';
# regrp_type = 1: 'MEM/OTHER/BUS'; =0:'CORP/UNF/GRP/GOV/OP/FENC'
hot1 <- hot1 %>% 
  filter(OTB!=0, days_prior<33) %>% 
  mutate(price=OTB_rev/OTB, OTB_cxl_rate = OTB_to_be_cxl/OTB,
         day_type = case_when(day_of_week == '7' ~ 1,
                              day_of_week == '1' ~ 1, 
                              day_of_week == '6' ~ 1, 
                              TRUE ~ 0),
         month = as.factor(month(stay_dt)),
         regrp_type = case_when(product_type == 'BUSINESS TRAVEL AGENCIES' ~ 1,
                                         product_type == 'MEMBERSHIP MARKETING' ~ 1,
                                         product_type == 'OTHER' ~ 1,
                                         TRUE ~ 0)) %>% 
  filter(!is.na(price), !is.na(daily_cxl_rev), !is.na(cummulative_cxl_rev))
 
h1_train <- hot1[hot1$stay_dt <= (max(hot1$stay_dt) - 21),]


h1_test <- hot1[hot1$stay_dt > (max(hot1$stay_dt) - 21),]

# For MASE, use the average survival (= average survival rate of all training data * OTB) as the naive forecast.
avg_svv_rt <- sum(h1_train$OTB_to_survive)/sum(h1_train$OTB)

nrow(h1_train)
nrow(h1_test)

```

##__KNN__
###__Variable selection 1__
```{r}
# variables:"days_prior", "day_type", "regrp_type", "price", "cummulative_gross_bookings""

keep_var = c("days_prior", "day_type", "regrp_type", "price", "cummulative_gross_bookings")


h1_trn_x <- as.data.frame(h1_train[, names(h1_train) %in% keep_var])
h1_trn_y <- as.data.frame(h1_train["OTB_cxl_rate"])

h1_tst_x <- as.data.frame(h1_test[, names(h1_test) %in% keep_var])
h1_tst_y <- as.data.frame(h1_test["OTB_cxl_rate"])

h1_train$predict_cxl_rate <- knn.reg(train = scale(h1_trn_x), test = scale(h1_trn_x), y = h1_trn_y, k = 20)$pred
h1_test$predict_cxl_rate <- knn.reg(train = scale(h1_trn_x), test = scale(h1_tst_x), y = h1_trn_y, k = 20)$pred
# k = 1, 2, ..., 40, 50, it seems like K=20 has generally better results. 

# In sample performance
eval(h1_train)
# Test dataset performance
eval(h1_test)
```

###__Variable selection 2__
```{r}
# variables:"days_prior", "day_type", "regrp_type", "price", "daily_cxl_rev""
# if replace daily_cxl_rev with daily_cxl_bookings, cummulative_gross_bookings, daily_net_bookings, daily_net_rev, cummulative_gross_rev, the results are similar but lower than this.
keep_var = c("days_prior", "day_type", "regrp_type", "price", "daily_cxl_rev")


h1_trn_x <- as.data.frame(h1_train[, names(h1_train) %in% keep_var])
h1_trn_y <- as.data.frame(h1_train["OTB_cxl_rate"])

h1_tst_x <- as.data.frame(h1_test[, names(h1_test) %in% keep_var])
h1_tst_y <- as.data.frame(h1_test["OTB_cxl_rate"])

h1_train$predict_cxl_rate <- knn.reg(train = scale(h1_trn_x), test = scale(h1_trn_x), y = h1_trn_y, k = 80)$pred
h1_test$predict_cxl_rate <- knn.reg(train = scale(h1_trn_x), test = scale(h1_tst_x), y = h1_trn_y, k = 80)$pred
# k = 1, 2, ..., 40, 50, it seems like K=80 has generally balanced results

# In sample performance
eval(h1_train)
# Test dataset performance
eval(h1_test)
```


###__Variable selection 3__
```{r}
# variables:"days_prior", "day_type", "regrp_type", "price", "cummulative_gross_bookings", "cummulative_cxl_rev"
# the combination of cum_gross_bk and cum_cxl_rev is better than cum_gross_bk and cum_cxl_bk
keep_var = c("days_prior", "day_type", "regrp_type", "price", "cummulative_gross_bookings", "cummulative_cxl_rev")

h1_trn_x <- as.data.frame(h1_train[, names(h1_train) %in% keep_var])
h1_trn_y <- as.data.frame(h1_train["OTB_cxl_rate"])

h1_tst_x <- as.data.frame(h1_test[, names(h1_test) %in% keep_var])
h1_tst_y <- as.data.frame(h1_test["OTB_cxl_rate"])

h1_train$predict_cxl_rate <- knn.reg(train = scale(h1_trn_x), test = scale(h1_trn_x), y = h1_trn_y, k = 50)$pred
h1_test$predict_cxl_rate <- knn.reg(train = scale(h1_trn_x), test = scale(h1_tst_x), y = h1_trn_y, k = 50)$pred
# k = 1, 2, ..., 40, 50, it seems like K=50 has generally balanced results

# In sample performance
eval(h1_train)
# Test dataset performance
eval(h1_test)
```



#__Hotel 2__
```{r}
summary(hot2)
```

##__Regroup day_of_week & product_type into 2-3 groups__
```{r}
# day_of_week cxl curves
hot2 %>% 
  filter(OTB!=0, days_prior<32, stay_dt <= max(stay_dt - 21)) %>% 
  group_by(days_prior, day_of_week) %>% 
  summarise(avg_cxl_rate = mean(OTB_to_be_cxl)/mean(OTB)) %>% 
  ggplot()+
  geom_line(aes(x=days_prior, y=avg_cxl_rate, color = day_of_week))+
  geom_dl(aes(label = day_of_week, x=days_prior, y=avg_cxl_rate, color = day_of_week), method = list(dl.combine("first.points", "last.points"), cex=0.5))

# product_type cxl curves
hot2 %>% 
  filter(OTB!=0, days_prior<32, stay_dt <= max(stay_dt - 21)) %>% 
  group_by(days_prior, product_type) %>% 
  summarise(avg_cxl_rate = mean(OTB_to_be_cxl)/mean(OTB)) %>% 
  ggplot()+
  geom_line(aes(x=days_prior, y=avg_cxl_rate, color = product_type))+
  geom_dl(aes(label = product_type, x=days_prior, y=avg_cxl_rate, color = product_type), method = list(dl.combine("first.points", "last.points"), cex=0.5))
```

+ According to the cxl curve, regroup the 7 days into 2 groups: 1,6,7 and 2,3,4,5

+ Regroup the product types into 2 groups:
    
    - group, tactical marketing, other, opaque, fenced
    - Goverment, business travel agencies, unfenced, corporate, membership marketing, wholesale
    
```{r}
# filter OTB=0, truncate the data by dp, add price, otb_cxl_rate, regrp_type, month, regrouping tags to the dataset
# day_type = 1: 'FRI/SAT/SUN'; = 0:'MON/TUE/WED/THU';
# regrp_type = 1: 'MEM/OTHER/BUS'; =0:'CORP/UNF/GRP/GOV/OP/FENC'
hot2 <- hot2 %>% 
  filter(OTB!=0, days_prior<33) %>% 
  mutate(price=OTB_rev/OTB, OTB_cxl_rate = OTB_to_be_cxl/OTB,
         day_type = case_when(day_of_week == '7' ~ 1,
                              day_of_week == '1' ~ 1, 
                              day_of_week == '6' ~ 1, 
                              TRUE ~ 0),
         month = as.factor(month(stay_dt)),
         regrp_type = case_when(product_type == 'GROUP' ~ 1,
                                         product_type == 'TACTICAL MARKETING' ~ 1,
                                         product_type == 'OTHER' ~ 1,
                                         product_type == 'OPAQUE' ~ 1,
                                         product_type == 'FENCED' ~ 1,
                                         TRUE ~ 0))
 
h2_train <- hot2[hot2$stay_dt <= (max(hot2$stay_dt) - 21),]


h2_test <- hot2[hot2$stay_dt > (max(hot2$stay_dt) - 21),]

# For MASE, use the average survival (= average survival rate of all training data * OTB) as the naive forecast.
avg_svv_rt <- sum(h2_train$OTB_to_survive)/sum(h2_train$OTB)

nrow(h2_train)
nrow(h2_test)
```

##__KNN__
###__Variable selection 1__
```{r}
# variables:"days_prior", "day_type", "regrp_type", "price", "cummulative_gross_bookings""

keep_var = c("days_prior", "day_type", "regrp_type", "price", "cummulative_gross_bookings")

h2_trn_x <- as.data.frame(h2_train[, names(h2_train) %in% keep_var])
h2_trn_y <- as.data.frame(h2_train["OTB_cxl_rate"])

h2_tst_x <- as.data.frame(h2_test[, names(h2_test) %in% keep_var])
h2_tst_y <- as.data.frame(h2_test["OTB_cxl_rate"])

h2_train$predict_cxl_rate <- knn.reg(train = scale(h2_trn_x), test = scale(h2_trn_x), y = h2_trn_y, k = 80)$pred
h2_test$predict_cxl_rate <- knn.reg(train = scale(h2_trn_x), test = scale(h2_tst_x), y = h2_trn_y, k = 80)$pred
# k = 1, 2, ..., 40, 50, it seems like K=80 has generally balanced results

# In sample performance
eval(h2_train)
# Test dataset performance
eval(h2_test)
```

###__Variable selection 2__
```{r}
# variables:"days_prior", "day_type", "regrp_type", "price", "daily_cxl_rev""
# if replace daily_cxl_rev with daily_cxl_bookings, cummulative_gross_bookings, daily_net_bookings, daily_net_rev, cummulative_gross_rev, the results are similar but lower than this.
keep_var = c("days_prior", "day_type", "regrp_type", "price", "cummulative_gross_rev")


h2_trn_x <- as.data.frame(h2_train[, names(h2_train) %in% keep_var])
h2_trn_y <- as.data.frame(h2_train["OTB_cxl_rate"])

h2_tst_x <- as.data.frame(h2_test[, names(h2_test) %in% keep_var])
h2_tst_y <- as.data.frame(h2_test["OTB_cxl_rate"])

h2_train$predict_cxl_rate <- knn.reg(train = scale(h2_trn_x), test = scale(h2_trn_x), y = h2_trn_y, k = 80)$pred
h2_test$predict_cxl_rate <- knn.reg(train = scale(h2_trn_x), test = scale(h2_tst_x), y = h2_trn_y, k = 80)$pred
# k = 1, 2, ..., 40, 50, it seems like K=80 has generally balanced results

# In sample performance
eval(h2_train)
# Test dataset performance
eval(h2_test)
```




