---
title: "Hotel Cancellation - Linear Regression Write UP"
author: "Caitlin Howansky & Wei Li"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: '3'
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r, include=FALSE}
# Run Linear Regression Rmd v2 first ###


# load package
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(tidyverse) 
library(dplyr) # joins
#library(janitor) # pretty cross-tabs
library(kableExtra) # pretty html tables
library(formattable)
library(gridExtra)
library(scales)
library(pastecs)
library(GGally)
library(lubridate)
library(directlabels)
library(Metrics)
library(tree)
library(ISLR)

load("my_env.RData")
```
##__Atlanta__
```{r, echo=FALSE}
tab_model(mod1_8)
left_join(left_join(mod1_8_mae,mod1_8_mape, by=c("dp_range")), mod1_8_mase, by=c("dp_range"))%>%
  arrange(meanAPE) %>%
  kable(col.names = c("Days Prior", "MAE", "MAPE", "MASE")) %>%
  kable_styling(bootstrap_options = c("condensed", "striped", "bordered")) %>%
  add_header_above(c("Atlanta Hotel" = 4))
```


For the Atlanta Hotel, the best model that was created took predicted the log of the cancellation rate. This worked best because the cancellation rate does not have a linear relationship with the dependent variables, but rather follows a more logarithmic function. It was also decided that some of the relationships between the cancellation rate and days prior was different by product type. Because of this, we introduced interaction terms of dummy variables and either days prior squared or took e to the power of days prior dependent on their cancellation curve. This model created an R-squared value of .5626. While these were the best results, it does not account for a lot of the data since cancellation rates of 0 cannot be included. When trying to account for this fact, the error numbers and R-squared value both got worse. This could be due to the unpredictability of having 0 cancellations in a day. While most of the variables in the model are statistically significant, many of the days of the week are not. This could be due to the fact that there is not a clear pattern for every day of the week. However, this model shows that Monday, Thursday, and Friday have higher cancellation rates than Sunday. From this model, you can see that one of the strongest variables influencing the cancellation rate is log of cummulative gross bookings. Another strong variable is the group type "group". We assumed this to be true as group has the largest amount of customers and therefore has the largest amount of bookings and cancellations. It is suprising how low the days prior coefficient is. This is most likely due to the fact that days prior is correlated with other variables which are taking on some of the effect of days prior.  This model, compared to the naive model performs relatively similar. However, the linear model performs slightly better in the date range between 15 - 21 days and 29 - 32 days according to the MASE number. This could be related to the volitility in the data. It is possible that the booking rates are more unpredictable further out, which is why smoothing the data is beneficial here. 


##__New York__
```{r, echo=FALSE}
tab_model(mod2_5)
left_join(left_join(mod2_5_mae,mod2_5_mape, by=c("dp_range")), mod2_5_mase, by=c("dp_range")) %>%
  arrange(meanAPE) %>%
  kable(col.names = c("Days Prior", "MAE", "MAPE", "MASE")) %>%
  kable_styling(bootstrap_options = c("condensed", "striped", "bordered")) %>%
  add_header_above(c("New York Hotel" = 4))
  
```

The New York hotel had better results from the linear regression models when compared to the Atlanta hotel. Like the Atlanta Hotel, the log of the cancellation rate was also predicted to to the nature of the relationship between the cancellation rate and days prior. For this model, the best grouping of product type was a group based off of similar average cancellation rates. The variables used in this model are the same as those that were used in the first hotel; however, dummy variable interaction terms for the product type grouping were not introduced. This is due to the fact that all but one of them followed a very logarithmic pattern. All of the interaction terms of days prior and the product types are significant except for the group including unfenced, wholesale, membership marketing, and business travel agencies. When comparing this model to the naive model, the naive model performs better in the first two date range groups, but the log liner regression performs better in the latter groups of dates. This is very similar to the way the Atlanta hotel behaved. Since it is harder to predict the cancellation rate when you are further away from the stay date, smoothing out predictions can help get better results. 


